---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


<!-- badges: start -->
<!-- badges: end -->

## Instruction
In 2007, the [original pilot project submission package](https://bitbucket.cdisc.org/projects/CED/repos/sdtm-adam-pilot-project/browse) was finalized and released following a review by FDA Staff, where the CDISC data and metadata contained within the package were evaluated for its suitability in meeting the needs and expectations of medical and statistical reviewers. In 2019, the [PHUSE Test Data Factory](https://advance.phuse.global/display/WEL/Test+Dataset+Factory) took on the goal of replicating the SDTM and ADaM data within the CDISC pilot package to match more modern data standards, bringing the ADaM data up to version 1.1. Atorus Research had regenerated the table outputs within the CDISC Pilot Project using the PHUSE Test Data Factory project’s data and the R Programming language. 

Our motivation behind this new project is to:
- Migrate Atorus project to a R package, convert the script to R functions.
- Pilot data from PHUSE Test Data Factory are convert and stored in this R package.
-	Demonstrate that we are able to obtain matching outputs using R
- Create a Shiny app to demostrate the tables.
-	Demonstrate Atorus's R package, `pharmaRTF`, in action.(You can find our package pharmaRTF right [here](https://github.com/atorus-research/pharmaRTF))


## The original data
To obtain the data for this repository, you can download the data from the PHUSE Github Repository, using [this link](https://github.com/phuse-org/phuse-scripts/blob/master/data/adam/TDF_ADaM_v1.0.zip) for ADaM data and [this link](https://github.com/phuse-org/phuse-scripts/blob/master/data/sdtm/TDF_SDTM_v1.0%20.zip)
 for the SDTM.

## Notes from Atorus project

-	The ADaMs we used to regenerate the CDISC Pilot displays were the PHUSE CDISC Pilot replication ADaMs following ADaM V1.1.  Since the CDISC Pilot displays were not regenerated using the PHUSE CDISC Pilot replication data there are likely discrepancies between the original CDISC Pilot analysis data and the PHUSE CDISC Pilot replication ADaMs.
-	SAS and R round differently.  While SAS rounds up if the value is 5 or greater, R rounds to the nearest even number.
-	In some circumstances, R packages will not produce a p-value if the the counts within the data are not high enough to make it statistically meaningful. 

## Notes on R Packages
As many programmers in the R community do, we relied on the [tidyverse](https://www.tidyverse.org/packages/) for much of our data processing. There are a few addition libraries that we used worth mentioning:
-	For the CMH test where testing for the alternate hypothesis that row means differ, the package vcdExtra was used, which is not included in the base distribution of R. We additionally had to make a slight modification to this library, which is available in [this fork]( https://github.com/mstackhouse/vcdExtra) of the package. The update is due to the fact that the `solve` function in R will throw an error when processing large, sparse tables. By replacing `solve` with `MASS::ginv`, the error is bypassed. We as an organization plan to perform further testing of this update and submit the update back to the original author.
-	For Mixed Models, the `lme4` package was used
-	For ANCOVA models, `car` was used and the `emmeans` package  was used to do LSMEANS.


## Installation

You can install the development version of datarx.pilotADaM from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("palandatarxcom/datarx.pilotADaM")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(datarx.pilotADaM)
## basic example code
```
